|

### 2021-07-20 nuova final draft.Rmd
titolo aggiustato + settati i seed

---
title: "Final draft"

@@

@@

  return(list(xs=xs,ws=ws,ess=ess))
}
```
We have already discussed the reasons why the SIS algorithm does not provide a good strategy in the filtering problem. We provide a graphical intuition of what happens when we use such filtering strategy on a simulated dataset. We decide to set $N=1000$,$m_{0}=0$,$C_{0}=100$ and $\tau=\sigma=1$. The results shown in the following two plots shows a clear degeneration of the effective sample size and bad fit of filtered states with respect to the true values.
```{r echo=FALSE}
N=1000
sis<-SISfun(y,N,m0,C0,tau,sigma)
sisxs<-sis$xs
sisws<-sis$ws
sisess<-sis$ess
sis.data<-data.frame(df,sisxs,sisws,sisess)

@@

    xs = rbind(xs,x)
    ws = rbind(ws,w)
    ess =rbind(ess,ESS)
  }
  return(list(xs=xs,ws=ws,ess=ess))
}
```
The estimated states of the Boostrap Particle Filter togheter with the effective sample size are shown in Figure XX. We decide to set $N=1000$,$m_{0}=0$,$C_{0}=100$ and $\tau=\sigma=1$. Notice how the resampling step allows the effective sample size not to drop, improving results.
```{r echo=FALSE, fig.align='center', fig.cap="Effective Sample Size", fig.height=4, fig.pos='H', fig.width=8, message=FALSE, warning=FALSE, results=F}
N=1000

@@

```{r include=FALSE}
truex<-x
estimatedx<-matrix(NA,ncol=4,nrow=6)
colnames(estimatedx)<-c("N","Threshold","KF","BPF")
estimatedx[,1]<-c(100,1000,10000,1000,1000,1000)
estimatedx[,2]<-c(0.5,0.5,0.5,0.5,0.25,0.1)

i=1
for(N in c(100,1000,10000)){

DLM1<-DLM(y,sigma2,tau2,m0,C0)
pf1<-PFfun(y,N,m0,C0,tau,sigma)

@@

i=4
for(k in c(2,5,10)){
  N=1000
DLM2<-DLM(y,sigma2,tau2,m0,C0)
pf2<-PFfun(y,N,m0,C0,tau,sigma,r=k)

@@

```{r echo=FALSE, fig.align='center', fig.cap="Comparison Bootstrap Particle Filter(BPF) and Kalman Filter (KF) for increasing number of generated particles (N)", fig.height=13, fig.pos='H', fig.width=10, message=FALSE, warning=FALSE, results=F}
N1=100
N2=1000
N3=10000

@@

```{r echo=FALSE}
gpf<-GPFfun(y,N,m0,C0,tau,sigma)
ESS.GPF<-ESSplot(df,gpf)
```

@@

```{r include=FALSE}
truex<-x
estimatedx<-matrix(NA,ncol=4,nrow=3)
colnames(estimatedx)<-c("N","Threshold","BPF","GPF")
estimatedx[,1]<-c(1000,1000,1000)
estimatedx[,2]<-c(0.5,0.25,0.1)

i=1
for(k in c(2,5,10)){
  N=1000

pf2<-PFfun(y,N,m0,C0,tau,sigma,r=k)
gpf2<-GPFfun(y,N,m0,C0,tau,sigma,r=k)

@@

```{r echo=FALSE, fig.align='center', fig.cap="Comparison Bootstrap Particle Filter(BPF) and Guided Particle Filter (GPF), number of generated particles N=1000", fig.height=4, fig.pos='H', fig.width=8, message=FALSE, warning=FALSE, results=F}
N2=1000

@@

```{r echo=FALSE}
apf<-APFfun(y,N,m0,C0,tau,sigma)
ESS.APF<-ESSplot(df,apf)
```

@@

```{r include=FALSE}
truex<-x
estimatedx<-matrix(NA,ncol=4,nrow=3)
colnames(estimatedx)<-c("N","Threshold","BPF","APF")
estimatedx[,1]<-c(1000,1000,1000)
estimatedx[,2]<-c(0.5,0.25,0.1)

i=1
for(k in c(2,5,10)){
  N=1000

pf2<-PFfun(y,N,m0,C0,tau,sigma,r=k)
apf2<-APFfun(y,N,m0,C0,tau,sigma,r=k)

@@

```{r echo=FALSE}
apfopt<-APFoptfun(y,N,m0,C0,tau,sigma)
ESS.APFopt<-ESSplot(df,apfopt)
```

@@

```{r echo=FALSE}
delta=0.7
lwf<-LWfun(y,N,m0,C0,1,1,1,1,delta,unif=T)
ESS.lwf<-ESSplot(df,lwf)
```

@@

>>>>

---
title: | 
 \Huge State-Space Models and Particle Filters

author:
  - Edoardo Marcelli
  - Roberto Colarieti
  - Alexei Verkhovtsev
  - Michele Bolla
  - Gabriele Romano
  - Benedetta Bruni

@@

date: "`r format(Sys.time(), '%d %B, %Y')`"

@@

  return(list(xs=xs,ws=ws,ess=ess))
}
```
We have already discussed the reasons why the SIS algorithm does not provide a good strategy in the filtering problem. We provide a graphical intuition of what happens when we use such filtering strategy on a simulated dataset. We decide to set $N=1000$,$m_{0}=0$,$C_{0}=100$ and $\tau=\sigma=1$. The results shown in the following two plots shows a clear degeneration of the effective sample size and bad fit of filtered states with respect to the true values.
```{r echo=FALSE}
set.seed(1991)
N=1000
sis<-SISfun(y,N,m0,C0,tau,sigma)
sisxs<-sis$xs
sisws<-sis$ws
sisess<-sis$ess
sis.data<-data.frame(df,sisxs,sisws,sisess)

@@

    xs = rbind(xs,x)
    ws = rbind(ws,w)
    ess =rbind(ess,ESS)
  }
  return(list(xs=xs,ws=ws,ess=ess))
}
```
The estimated states of the Boostrap Particle Filter togheter with the effective sample size are shown in Figure XX. We decide to set $N=1000$,$m_{0}=0$,$C_{0}=100$ and $\tau=\sigma=1$. Notice how the resampling step allows the effective sample size not to drop, improving results.
```{r echo=FALSE, fig.align='center', fig.cap="Effective Sample Size", fig.height=4, fig.pos='H', fig.width=8, message=FALSE, warning=FALSE, results=F}
set.seed(1991)
N=1000

@@

```{r include=FALSE}
set.seed(1991)
truex<-x
estimatedx<-matrix(NA,ncol=4,nrow=6)
colnames(estimatedx)<-c("N","Threshold","KF","BPF")
estimatedx[,1]<-c(100,1000,10000,1000,1000,1000)
estimatedx[,2]<-c(0.5,0.5,0.5,0.5,0.25,0.1)

i=1
for(N in c(100,1000,10000)){
set.seed(1991)
DLM1<-DLM(y,sigma2,tau2,m0,C0)
pf1<-PFfun(y,N,m0,C0,tau,sigma)

@@

i=4
for(k in c(2,5,10)){
  N=1000
  set.seed(1991)
DLM2<-DLM(y,sigma2,tau2,m0,C0)
pf2<-PFfun(y,N,m0,C0,tau,sigma,r=k)

@@

```{r echo=FALSE, fig.align='center', fig.cap="Comparison Bootstrap Particle Filter(BPF) and Kalman Filter (KF) for increasing number of generated particles (N)", fig.height=13, fig.pos='H', fig.width=10, message=FALSE, warning=FALSE, results=F}
set.seed(1991)
N1=100
N2=1000
N3=10000

@@

```{r echo=FALSE}
set.seed(1991)
gpf<-GPFfun(y,N,m0,C0,tau,sigma)
ESS.GPF<-ESSplot(df,gpf)
```

@@

```{r include=FALSE}
set.seed(1991)
truex<-x
estimatedx<-matrix(NA,ncol=4,nrow=3)
colnames(estimatedx)<-c("N","Threshold","BPF","GPF")
estimatedx[,1]<-c(1000,1000,1000)
estimatedx[,2]<-c(0.5,0.25,0.1)

i=1
for(k in c(2,5,10)){
  N=1000
set.seed(1991)
pf2<-PFfun(y,N,m0,C0,tau,sigma,r=k)
gpf2<-GPFfun(y,N,m0,C0,tau,sigma,r=k)

@@

```{r echo=FALSE, fig.align='center', fig.cap="Comparison Bootstrap Particle Filter(BPF) and Guided Particle Filter (GPF), number of generated particles N=1000", fig.height=4, fig.pos='H', fig.width=8, message=FALSE, warning=FALSE, results=F}
set.seed(1991)
N2=1000

@@

```{r echo=FALSE}
set.seed(1991)
apf<-APFfun(y,N,m0,C0,tau,sigma)
ESS.APF<-ESSplot(df,apf)
```

@@

```{r include=FALSE}
set.seed(1991)
truex<-x
estimatedx<-matrix(NA,ncol=4,nrow=3)
colnames(estimatedx)<-c("N","Threshold","BPF","APF")
estimatedx[,1]<-c(1000,1000,1000)
estimatedx[,2]<-c(0.5,0.25,0.1)

i=1
for(k in c(2,5,10)){
  N=1000
set.seed(1991)
pf2<-PFfun(y,N,m0,C0,tau,sigma,r=k)
apf2<-APFfun(y,N,m0,C0,tau,sigma,r=k)

@@

```{r echo=FALSE}
set.seed(1991)
apfopt<-APFoptfun(y,N,m0,C0,tau,sigma)
ESS.APFopt<-ESSplot(df,apfopt)
```

@@

```{r echo=FALSE}
delta=0.7
set.seed(1991)
lwf<-LWfun(y,N,m0,C0,1,1,1,1,delta,unif=T)
ESS.lwf<-ESSplot(df,lwf)
```

@@

Finally, we compare the performances of the Liu and West filter and the other filtering strategies discussed so far. As we expect, in the linear Gaussian case, Kalman filter outperforms the other strategies. On the other hand, the Liu and West, with hyperparameters setted as described above, provides the worst estimates in terms of RMSE. This is not necessarily due to an overall weakness of the LWF with respect to the other filters, but simply to a not accurate choice of the hyperparameters. Moreover, while increasing the number of particles improves the accuracy, notice that the threshold for ESS-based resampling is not strictly related to better performances.

```{r include=FALSE}
set.seed(1991)
truex<-x
estimatedx<-matrix(NA,ncol=6,nrow=6)
colnames(estimatedx)<-c("N","Threshold","KF","BPF","APF","LWF")
estimatedx[,1]<-c(100,1000,10000,1000,1000,1000)
estimatedx[,2]<-c(0.5,0.5,0.5,0.5,0.25,0.1)

i=1
for(N in c(100,1000,10000)){
set.seed(1991)
DLM1<-DLM(y,sigma2,tau2,m0,C0)
pf1<-PFfun(y,N,m0,C0,tau,sigma)
apf1<-APFfun(y,N,m0,C0,tau,sigma)
lwf1<-LWfun(y,N,m0,C0,1,1,1,1,delta,unif=T)

estimatedx[i,3]<-RMSE(truex,DLM1$m)
estimatedx[i,4]<-RMSE(truex,Filtervalues(pf1)$mean)
estimatedx[i,5]<-RMSE(truex,Filtervalues(apf1)$mean)
estimatedx[i,6]<-RMSE(truex,Filtervalues(lwf1)$mean)
i=i+1

i=4
for(k in c(2,5,10)){
  set.seed(1991)
  N=10
DLM2<-DLM(y,sigma2,tau2,m0,C0)
pf2<-PFfun(y,N,m0,C0,tau,sigma,r=k)
apf2<-APFfun(y,N,m0,C0,tau,sigma,r=k)
lwf2<-LWfun(y,N,m0,C0,1,1,1,1,delta,unif=T,r=k)

estimatedx[i,3]<-RMSE(truex,DLM2$m)
estimatedx[i,4]<-RMSE(truex,Filtervalues(pf2)$mean)
estimatedx[i,5]<-RMSE(truex,Filtervalues(apf2)$mean)
estimatedx[i,6]<-RMSE(truex,Filtervalues(lwf2)$mean)
i=i+1

```{r echo=FALSE, message=FALSE, warning=FALSE, results="asis"}
kbl(estimatedx, align="c", longtable = T, booktabs = T, caption="RMSE",digits=3)
```
